Measurement and Probability in Fuzzy Quantum Theories

### 2017-03-13 prop_letter.tex
Clean up files...

\begin{document}
\title{Measurement and Probability in Fuzzy Quantum
Theories}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Textbook quantum mechanics asserts that, given a quantum spin system
described by the wave function
$\ket{\psi}=\frac{\sqrt{\pi}}{2}\fket{\ps\frac{1}{2}}+\frac{\sqrt{4-\pi}}{2}\fket{\minus\frac{1}{2}}$
and given the observable
$S_z =
\frac{\hbar}{2}(\fproj{\ps\frac{1}{2}}-\fproj{\minus\frac{1}{2}})$,
then the probability of observing the eigenvalue $\ps\frac{\hbar}{2}$
(with post-measurement state $\fket{\ps\frac{1}{2}}$) is
$\frac{\pi}{4}$ and the probability of observing the eigenvalue
$\minus\frac{\hbar}{2}$ (with post-measurement state
$\fket{\minus\frac{1}{2}}$) is $1-\frac{\pi}{4}$. According to the
frequency interpretation of probability, the above assertion implies
that when performing a very large number of experiments on quantum
systems described by wave function $\ket{\psi}$, a fraction
$\frac{\pi}{4}$ of the experiments will result in the observation
$\ps\frac{\hbar}{2}$ and the remaining fraction $1-\frac{\pi}{4}$ will
result in the observation $\minus\frac{\hbar}{2}$.

In reality, i.e., in an actual experiment, simulation, or computation,
bounded by space, time, and energy resources, the situation is much
fuzzier: (i) we cannot \emph{exactly} prepare the state $\ket{\psi}$;
(ii) we cannot build a measurement apparatus corresponding
\emph{exactly} to observable $S_z$; and (iii) we cannot prepare
\emph{exact} replicas of any state. In other words, we can only
perform a limited number of experiments on quantum systems closely
related to wave function $\ket{\psi}$ and using an apparatus that
closely approximates observable $S_z$. Given any fixed resources,
these experiments will \emph{not} result in the \emph{exact} fraction
$\frac{\pi}{4}$ of observations $\ps\frac{\hbar}{2}$ and
$1-\frac{\pi}{4}$ of observations $\minus\frac{\hbar}{2}$. In fact,
these experiments can be viewed as a method (algorithm) to calculate
the value $\pi$ and it is known that the state of the art algorithms
for computing the $n$th binary digit of~$\pi$ require on the order of
$O\left(n\log^{O\left(1\right)}\left(n\right)\right)$
operations~\cite{journals/moc/BaileyBP97}. In other words, calculating
the $n$th digit of $\pi$ appears to require more and more physical
resources as~$n$ gets larger.

All of this is well-known and holds in both the classical and quantum
worlds. In the classical world, these observations are believed to be
well-understood approximations to an idealized independent reality
that can continuously approached with more and more resources. The
question in the quantum world is more subtle; it is not evident at all
that this resource-aware perspective does not alter the very
foundations of quantum mechanics. Indeed taking such resource bounds
into consideration is what founded computer science as a discipline
and has become crucial for understanding the very nature of
computation. Following Feynman~\cite{Feynman1982Simulating},
Landauer~\cite{Landauer1996188}, and others, one might argue that this
resource-bounded perspective is also crucial for understanding the
very nature of physical processes.

In previous work, we analyzed the wave function description of quantum
systems from a resource-bounded perspective. Technically, instead of
postulating that quantum states are rays in a Hilbert space, i.e., we
instead used special finite fields whose size reflected the available
resources. \ldots

\newpage

This is a \emph{theoretical} investigation of \emph{experimental}
physics using \emph{computational} methods. All experiments and computations
are processes bounded in space, time, energy, and other resources~\cite{Jaeger2007,Piccinini2015}.
Yet, for centuries, the mathematical formalization of such processes
has been founded on the infinitely precise real or complex numbers~\cite{Ziegler2007,weihrauch2012computable,blum2012complexity}.
Our purpose here is to exploit the consequences of replacing infinitely-specified
quantum probabilities by finite number of intervals used in interval-valued
probability measures~(IVPM); in particular, we show that a quantum
IVPM may not be induced by a infinitely-precise state, but it might
be more and more likely to be identified to a particular state when
the measurement resources increase.

Indeed, almost every description of quantum mechanics, quantum computation,
or quantum experiments refers to entities such as $\rme$, $\pi$,
$\sqrt{2}$, etc (see, e.g., \cite{Redhead1987-REDINA,544199,Mermin2007}).
From a computational perspective, such numbers do not exist in their
entirety ``for free~\cite{Kent1999,CliftonKent2000}.'' For example,
the state of the art algorithms for computing the $n$th binary digit
of~$\pi$ require on the order of $O\left(n\log^{O\left(1\right)}\left(n\right)\right)$
operations~\cite{journals/moc/BaileyBP97}. In other words, simply
referring to the $n$th digit of $\pi$ requires more and more resources
as $n$ gets larger. Taking such resource bounds into consideration
is what founded computer science as a discipline and is crucial for
understanding the very nature of computation and, following Feynman~\cite{Feynman1982Simulating},
Landauer~\cite{Landauer1996188}, and others, for understanding the
very nature of physical processes.

We have been revisiting quantum mechanics, quantum information, and
quantum computation from this resource-aware perspective. Our initial
results in that domain showed how subtle the issues can be~\cite{usat,geometry2013,DQT2014}:
a straightforward replacement of the complex numbers by a finite field
yields a variant of quantum mechanics in which computationally hard
problems like UNIQUE-SAT (which decides whether a given Boolean formula
is unsatisfiable or has exactly one satisfying assignment~\cite{Valiant198685,Papadimitriou1993,AroraBarak2009})
can be deterministically solved in constant time. To eliminate such
unrealistic theories requires delicate analysis of the structure of
the Hilbert space, the process of observation, and the notion of probability
teasing apart their reliance on the infinitely precise real numbers~\cite{geometry2013,DQT2014}.

In this paper, we shift focus from the infinitely-specified but not
directly observable quantum states, to observable measurable properties
of quantum systems and their probabilities. Furthermore, we insist
that our theories of measurement and probability only refer to finitely
communicable evidence within feasible computational bounds. It follows
that states, observations, and probabilities all become ``fuzzy'',
i.e., specified by intervals of confidence that can only increase
in precision if the available resources increase proportionally. Our
notion of ``fuzzy quantum mechanics'' is related to existing work~\cite{GranikCaulfield1996,Pykacz2013,SNL2009,Gudder2005,aerts1993physical}
but, as will be explained in more detail, is distinguished by its
unique computational character.

We will begin by reviewing existing work that recasts classical probability
spaces in a resource-aware setting and move to aim at recasting quantum
probability and quantum measurement. In particular, we develop a measurement
framework based on quantum IVPM. Surprisingly, we found a quantum
IVPM which can not be induced by a state, while Shapley proved a classical
convex IVPM can always be induced by a classical ``state''~\cite{Shapley1971,Grabisch2016},
and Gleason proved a infinitely-specified quantum probabilities measure
can always be induced by a quantum state~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.
Nevertheless, a quantum IVPM could still correspond to many possible
states if it is broken into pieces. Our examples also suggest that
a quantum IVPM could be more and more likely induced by a state while
the intervals become sharper, i.e., the measurement resources increase.

Since Gleason's theorem supports the idea of von Neumann to define
a state as a density matrix~\cite{Varadarajan2008}, a density matrix
might not be a adequate imprecise quantum state. Instead, a imprecise
quantum state might be more like an interactive system suggested by
Quantum Bayesianism or QBism~\cite{FuchsSchack2013,FuchsMerminSchack2014,VonBaeyer2016},
which would interact differently with different clients. Furthermore,
the validity the fundamental theorems of quantum mechanics followed
up by Gleason's theorem, such as Bell~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and Kochen-Specker~\cite{kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007},
might need to be reassessed based on our imprecise measurement. Our
research then might provide a new insight to the old debate among
Meyer, Mermin, and others about whether finite precision measurement
would nullify the Kochen-Specker theorem or not~\cite{PhysRevLett.83.3751,Mermin1999,BarrettKent2004}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Classical Probability}

A \emph{probability space} specifies the necessary conditions for
reasoning coherently about collections of uncertain
events~\cite{Kolmogorov1950,Shafer1976,Griffiths2003,Swart2013}.  We
review the conventional presentation of probability spaces and then
discuss the computational resources needed to estimate probabilities.

%%%
\subsection{Classical Probability Spaces}

The conventional definition of a probability space builds upon the
field of real numbers. In more detail, a probability space consists
of a \emph{sample space} $\Omega$, a space of \emph{events}~$\events$,
and a \emph{probability measure}~$\pmeas$ mapping events in $\events$
to the real interval $[0,1]$. We will only consider \emph{finite}
sets of events and restrict our attention to non-empty finite sets
$\Omega$ as the sample space. The space of events $\events$ includes
every possible subset of $\Omega$: it is the powerset~$2^{\Omega}=\set{E}{E\subseteq\Omega}$.
For future reference, we emphasize that events are the primary notion
of interest and that the sample space is a convenient artifact that
allows us to treat events as sets obeying the laws of Boolean algebra~\cite{Boole1948,Redhead1987-REDINA,Griffiths2003}.

\begin{definition}[Probability Measure]\label{def:ClassicalProbabilitySpace}
  Given the set of events $\events$, a \emph{probability measure} is a
  function $\pmeas:\events\rightarrow[0,1]$ such that:
\begin{itemize}
\item $\pmeas(\emptyset)=0$,
\item $\pmeas(\Omega)=1$, 
\item for every event $E$,
  $\pmeas\left(\Omega\backslash E\right)=1-\pmeas\left(E\right)$ where
  $\Omega\backslash E$ is the complement event of $E$, and
\item for every collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of
  pairwise disjoint events,
  $\pmeas\left(\bigcup_{i=1}^{N}E_{i}\right)=\sum_{i=1}^{N}\pmeas(E_{i})$.
\end{itemize}
\end{definition}
\noindent There is some redundancy in the definition that will be useful when
moving to quantum probability spaces.

\begin{example}[Two-coins experiment]\label{ex1} Consider an
  experiment that tosses two coins. We have four possible outcomes
  that constitute the sample space $\Omega=\{HH,HT,TH,TT\}$. There are
  16 total events including the event $\{HH,HT\}$ that the first coin
  lands heads up, the event $\{HT,TH\}$ that the two coins land on
  opposite sides, and the event $\{HT,TH,TT\}$ that at least one coin
  lands tails up. Here is a possible probability measure for these
  events:
\begin{equation}
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) & = & 0\\
\pmeas(\{HH\}) & = & \frac{1}{3}\\
\pmeas(\{HT\}) & = & 0\\
\pmeas(\{TH\}) & = & \frac{2}{3}\\
\pmeas(\{TT\}) & = & 0\\
\pmeas(\{HH,HT\}) & = & \frac{1}{3}\\
\pmeas(\{HH,TH\}) & = & 1\\
\pmeas(\{HH,TT\}) & = & \frac{1}{3}
\end{array} & \begin{array}{rcl}
\pmeas(\{HT,TH\}) & = & \frac{2}{3}\\
\pmeas(\{HT,TT\}) & = & 0\\
\pmeas(\{TH,TT\}) & = & \frac{2}{3}\\
\pmeas(\{HH,HT,TH\}) & = & 1\\
\pmeas(\{HH,HT,TT\}) & = & \frac{1}{3}\\
\pmeas(\{HH,TH,TT\}) & = & 1\\
\pmeas(\{HT,TH,TT\}) & = & \frac{2}{3}\\
\pmeas(\{HH,HT,TH,TT\}) & = & 1
\end{array}\end{array}
\end{equation}
\end{example}

\noindent It is useful to think that this probability measure is
completely determined by the ``reality'' of the two coins in question
and their characteristics, in the sense that each pair of coins
induces a measure, and each measure must correspond to some pair of
coins. The measure above would be induced by two particular coins such
that the first coin is twice as likely to land tails up than heads up
and the second coin is double-headed.

%%%%%
\subsection{Measuring Probabilities: Buffon's Needle Problem\label{subsec:Measuring-Probabilities:-Buffon}}

In a strict computational or experimental setting, one should question
the reliance of the definition of probability space on the uncountable
and uncomputable real
interval~$[0,1]$~\cite{Turing_1937,Ziegler2007,weihrauch2012computable}.
This interval includes numbers like $0.h_{1}h_{2}h_{3}\ldots$ where
$h_{i}$ is 1 or 0 depending on whether Turing machine
$\mathit{TM}_{i}$ halts or not. Such numbers cannot be computed. This
interval also includes numbers like $\pi / 4$ which can only be
computed with increasingly large resources as the precision
increases. Therefore, in a resource-aware computational or
experimental setting, it is more appropriate to consider probability
measures that map events to a set of elements computable with a fixed
set of resources. We expand on this observation in the next section
and then recall its formalization using interval-valued probability
measures~\cite{Weichselberger2000,JamisonLodwick2004}.

In the previous example, we assumed the probability~$\pmeas(E)$ of
each event~$E$ is known a priori. In reality, although each event is
assumed to have a probability, the exact value of $\pmeas(E)$ may not
be known. According to the \emph{frequency interpretation of
  probability} (which we will revisit when moving to the quantum
case)~\cite{Venn1876,Hajek2012}, 
to determine the probability of an event, we run $M$
independent trials which gives us an approximation of the (assumed)
``true'' or ``real'' probability. Let $x_{i}$ be 1 or 0 depending on
whether the event~$E$ occurs in the $i$-th trial or not, then
$\pmeas(E)$ could be approximated to given accuracy~$\epsilon>0$ by
the relative frequency~$\sum_{i=1}^{M}x_{i} / M$ with the
probability converging to one as $M$ goes to infinity, i.e.,
\begin{equation}
\forall\epsilon>0,\lim_{M\rightarrow\infty}\pmeas\left(\left|\pmeas(E)-\frac{1}{M}\sum_{i=1}^{M} x_{i}\right|<\epsilon\right)=1\textrm{ .}
\end{equation}
This fact is called the law of large numbers~\cite{Bernoulli2006,Kolmogorov1950,Uspensky1937,Shafer1976,544199}.

Let's look at a concrete example. Suppose we drop a needle of length
$\ell$ onto a floor made of equally spaced parallel lines a distance
$h$ apart, where $\ell<h$. It is a known fact that the probability of
the needle crossing a line is
$2\ell/\left(\pi h\right)$~\cite{Buffon1777,DeMorgan1872,Hall1873,Uspensky1937}.
Consider an experimental setup consisting of a collection of $M$
identical needles of length $\ell$. We throw the $M$ needles one
needle at a time, and observe the number $M_c$ of needles that cross a
line, thus estimating the probability of a needle crossing a line to
be $M_c/M$. In an actual experiment with $500$ needles and the
ratio $\ell/h=0.75$~\cite{Hall1873}, it was found that $236$
crossed a line so the relative frequency is $0.472$ whereas the
idealized mathematical probability is $0.4774\ldots$.  In a larger
experiment with $5000$ needles and the ratio
$\ell/h=0.8$~\cite{Uspensky1937}, the relative frequency was
calculated to be $0.5064$ whereas the idealized mathematical
probability is $0.5092\ldots$. We see that the observed probability
approaches $2\ell/\left(\pi h\right)$ but only if \emph{larger and larger
  resources} are expended. These resource considerations suggest that
it is possible to replace the real interval $[0,1]$ with rational
numbers up to a certain precision related to the particular experiment
in question. There is clearly a connection between the number of
needles and the achievable precision: in the hypothetical experiment
with 3 needles, it is not sensible to retain 100 digits in the
expansion of $2\ell/\left(\pi h\right)$.

There is another more subtle assumption of unbounded computational
power in the experiment. We are assuming that we can always determine
with certainty whether a needle is crossing a line. But ``lines'' on
the the floor have thickness, their distance apart is not exactly $h$,
and the needles' lengths are not all absolutely equal to $\ell$.
These perturbations make the events ``fuzzy.'' Thus, in an experiment
with limited resources, it is not possible to talk about the idealized
event that exactly $M_c$ needles cross lines as this would require the
most expensive needles built to the most precise accuracy, laser
precision for drawing lines on the floor, and the most powerful
microscopes to determine if a needle does cross a line. Instead we
might talk about the event that $M_c-\delta$ needles evidently cross
lines and $M_c+\delta'$ needles plausibly cross lines where $\delta$ and
$\delta'$ are resource-dependent approximations. This fuzzy notion of
events leads to probabilities being only calculable within intervals
of confidence reflecting the certainty of events and their
plausibility. This is indeed consistent with published experiments: in
an experiment with $3204$ needles and the ratio
$\ell/h=0.6$~\cite{DeMorgan1872}, $1213$ needles clearly
crossed a line and $11$ needles were close enough to plausibly be
considered as crossing the line: we would express the probability in
this case as the interval
$\left[\frac{1213}{3204},\frac{1224}{3204}\right]$ expressing that we
are certain that the event has probability at least
$\frac{1213}{3204}$ but it is possible that it would have probability
$\frac{1224}{3204}$.

Another way to re-express the above observations is that the calculus
of probabilities requires all probabilities to be \emph{coherent}
(e.g., the probability of the union of two disjoint events is the sum
of their probabilities). But given limited resources, the calculation
of a probability measure can only be approximate and hence the
probabilities cannot be absolutely guaranteed to be coherent in the
ideal sense.

%%%%%
\subsection{Classical Interval-Valued Probability Measures}

As motivated above, an event $E_{1}$ may have an interval of probability
$[l_{1},r_{1}]$. Assume that another disjoint event $E_{2}$ has
an interval of probability $[l_{2},r_{2}]$, what is the interval
of probability for the event $E_{1}\cup E_{2}$? The answer is somewhat
subtle: although it is possible to use the sum of the intervals $[l_{1}+l_{2},r_{1}+r_{2}]$
as the combined probability, one can find a much tighter interval
if information \emph{against} the event (i.e., information about the
complement event) is also taken into consideration. Formally, for
a general event $E$ with probability $[l,r]$, the evidence that
contradicts $E$ is evidence supporting the complement of $E$. The
complement of $E$ must therefore have probability $\left[1-r,1-l\right]$
which we abbreviate $\left[1,1\right]-\left[l,r\right]$. Given a
sample space~$\Omega$, its set of events~$\events$, and a collection
of intervals $\mathscr{I}$, a function~$\bar{\mu}:\events\rightarrow\mathscr{I}$
is a classical interval-valued probability measure if and only if
$\bar{\mu}$ satisfies the following conditions~\cite{JamisonLodwick2004}
where the last line uses $\subseteq$ to allow for tighter intervals
that exploit the complement event: 
\begin{itemize}
\item $\bar{\mu}(\emptyset)=[0,0]$. 
\item $\bar{\mu}(\Omega)=[1,1]$. 
\item For any event $E$, $\bar{\mu}\left(\Omega\backslash
E\right)=\left[1,1\right]-\bar{\mu}\left(E\right)$.
\item For a collection $\left\{ E_{i}\right\} _{i=1}^{N}$ of pairwise disjoint
events, we have 
\begin{equation}
\bar{\mu}\left(\bigcup_{i=1}^{N}E_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(E_{i}\right)\textrm{ .}\label{eq:classical-include}
\end{equation}
\end{itemize}

\begin{example}[Two-coin experiment with interval probability]
\label{ex3} We split the unit interval $[0,1]$ in the following
four closed sub-intervals: $[0,0]$ which we call \imposs, $[0,\frac{1}{2}]$
which we call \unlikely, $[\frac{1}{2},1]$ which we call \likely,
and $[1,1]$ which we call \necess. Using these new values, we can
modify the probability measure of example~\ref{ex1} by mapping each
numeric value to the smallest sub-interval containing it to get the
following: 
\begin{equation}
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\bar{\mu}(\emptyset) & = & \imposs\\
\bar{\mu}(\{HH\}) & = & \unlikely\\
\bar{\mu}(\{HT\}) & = & \imposs\\
\bar{\mu}(\{TH\}) & = & \likely\\
\bar{\mu}(\{TT\}) & = & \imposs\\
\bar{\mu}(\{HH,HT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH\}) & = & \necess\\
\bar{\mu}(\{HH,TT\}) & = & \unlikely
\end{array} & \begin{array}{rcl}
\bar{\mu}(\{HT,TH\}) & = & \likely\\
\bar{\mu}(\{HT,TT\}) & = & \imposs\\
\bar{\mu}(\{TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH\}) & = & \necess\\
\bar{\mu}(\{HH,HT,TT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH,TT\}) & = & \necess\\
\bar{\mu}(\{HT,TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH,TT\}) & = & \necess
\end{array}\end{array}
\end{equation}
Despite the absence of infinitely precise numeric information, the
probability measure is quite informative: it reveals that the second
coin is double-headed and that the first coin is biased. To understand
the $\subseteq$-condition, consider the following calculation:
\begin{equation}\eqalign{
 & \bar{\mu}(\{HH\})+\bar{\mu}(\{HT\})+\bar{\mu}(\{TH\})+\bar{\mu}(\{TT\})\\
=& \imposs+\unlikely+\imposs+\likely\\
=& \left[0,0\right]+\left[0,\frac{1}{2}\right]+\left[0,0\right]+\left[\frac{1}{2},1\right]=
\left[\frac{1}{2},\frac{3}{2}\right]\textrm{ .}
}\end{equation}
If we were to equate $\bar{\mu}(\Omega)$ with the sum of the individual
probabilities, we would get that $\bar{\mu}(\Omega)=\left[\frac{1}{2},\frac{3}{2}\right]$.
However, using the fact that $\bar{\mu}(\emptyset)=\imposs$, we have
$\bar{\mu}\left(\Omega\right)=1-\bar{\mu}\left(\emptyset\right)=\necess=[1,1]$.
This interval is tighter and a better estimate for the probability
of the event $\Omega$, and of course it is contained in $[\frac{1}{2},\frac{3}{2}]$.
However it is only possible to exploit the information about the complement
when all four events are combined. Thus the $\subseteq$-condition
allows us to get an estimate for the combined event from each of its
constituents and then gather more evidence knowing the aggregate
event.
\end{example}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Quantum Probability}

The mathematical framework of classical probability above assumes that
there exists a predetermined set of events that are independent of the
particular experiment --- classical physics is
non-contextual~\cite{kochenspecker1967,Redhead1987-REDINA,peres1995quantum,Jaeger2007}. 
However, even in classical situations, the structure
of the event space is often only partially known and the precise
dependence of two events on each other cannot, a priori, be determined
with certainty. In the quantum framework, this partial knowledge is
compounded by the fact that there exist non-commuting events which
cannot happen simultaneously. To accommodate these more complex
situations, conventional approaches to quantum probability abandon the
sample space~$\Omega$ and reason directly about events which are
generalized from plain sets to projection operators. A quantum
probability space therefore consists of just two components: a set of
events $\qevents$ often formalized as projection operators and a
probability measure $\mu:\qevents\rightarrow[0,1]$ formalized using
the Born rule~\cite{Born1983,Mermin2007,Jaeger2007}.

%%%%%
\subsection{Quantum Events}

\begin{definition}[Projection Operators; Orthogonality~\cite{10.2307/2308516,Redhead1987-REDINA,peres1995quantum,Griffiths2003,Swart2013}]
  \label{def:Projection} Given a Hilbert space $\Hilb$, an event (an
  experimental proposition~\cite{BirkhoffVonNeumann1936}, a
  question~\cite{10.2307/2308516,Abramsky2012},
  or an elementary quantum test~\cite{peres1995quantum}) is represented
  as a (self-adjoint or orthogonal~\cite {Griffiths2003,Maassen2010})
  projection operator $P:\Hilb\rightarrow\Hilb$ onto a linear subspace
  of $\Hilb$. The following define projections and list some of their properties:
\begin{itemize}
\item $\mathbb{0}$ is a projection. 
\item For any pure state~$\ket{\psi}$, $\proj{\psi}$ is a projection
operator. 
\item Projection operators $P_{0}$ and $P_{1}$ are \emph{orthogonal}
  if $P_{0}P_{1}=P_{1}P_{0}=\mathbb{0}$. The sum of two projection
  operators~$P_{0}+P_{1}$ is also a projection operator if and only if
  they are orthogonal.
\item Conversely, every projection~$P$ can be expressed as
  $\sum_{j=1}^{N}\proj{\psi_{j}}$, where $P$ actually projects onto
  the linear subspace with orthonormal
  basis~$\left\{ \ket{\psi_{j}}\right\} _{j=1}^{N}$.
\item A set of projections $\left\{ P_{i}\right\} _{i=1}^{N}$ is
  called an \emph{ideal measurement} if it is a partition of the
  identity, i.e., $\sum_{i=1}^{N}P_{i}=\mathbb{1}$~\cite{Swart2013}.
  In this case, projections $\left\{ P_{i}\right\} _{i=1}^{N}$ must
  be mutually orthogonal~\cite{Griffiths2003,Halmos1957}, and $N$ must be less
  or equal to the dimension of the Hilbert space.
\item If $P$ is a projection operator, then $\mathbb{1}-P$ is also a
  projection operator, called its \emph{complement}. It is orthogonal to
  $P$, and corresponds to the complement event~$\Omega\backslash E$ in
  classical probability~\cite{Griffiths2003}.
\item Projection operators $P_{0}$ and $P_{1}$ \emph{commute} if $P_{0}P_{1}=P_{1}P_{0}$.
The product of two projection operators~$P_{0}P_{1}$ is also a projection
operator if and only if they commute. This corresponds to the classical
intersection between events~\cite{peres1995quantum,Griffiths2003}. 
\item For two commuting projection operators $P_{0}$ and $P_{1}$,
  their \emph{disjunction}~$P_{0}\vee P_{1}$ is defined to be
  $P_{0}+P_{1}-P_{0}P_{1}$~\cite{Griffiths2003}.
\end{itemize}
\end{definition}

\begin{example}[One-qubit quantum probability space] Consider a
  one-qubit Hilbert space with each event interpreted as a possible
  post-measurement state~\cite{peres1995quantum,Mermin2007,Jaeger2007}. 
  For example, the event $\proj{0}$ indicates
  that the post-measurement state will be $\ket{0}$; the probability
  of such an event depends on the current state; the event $\proj{1}$
  indicates that the post-measurement state will be $\ket{1}$; the
  event $\proj{\ps}$ where
  $\ket{\ps}=\left(\ket{0}+\ket{1}\right) / \sqrt{2}$ indicates that the
  post-measurement state will be $\ket{\ps}$; the event
  $\mathbb{1}=\proj{0}+\proj{1}$ indicates that the post-measurement
  state will be a linear combination of $\ket{0}$ and $\ket{1}$ and is
  clearly certain; finally the empty event $\mathbb{0}$ states that
  the post-measurement state will be the empty state and is
  impossible. As in the classical case, a probability measure is a
  function that maps events to $[0,1]$. Here is a partial
  specification of a possible probability measure that would be
  induced by a system whose current state is $\ket{0}$,
  $\mu\left(\mathbb{0}\right)=0$, $\mu\left(\mathbb{1}\right)=1$,
  $\mu\left(\proj{0}\right)=1$, $\mu\left(\proj{1}\right)=0$,
  $\mu\left(\proj{\ps}\right)=\frac{1}{2}$, \ldots. Note that, similarly to
  the classical case, the probability of $\mathbb{1}$ is 1 and the
  probability of collections of orthogonal events (e.g.,
  $\proj{0}+\proj{1}$) is the sum of the individual probabilities.  A
  collection of non-orthogonal events (e.g., $\proj{0}$ and
  $\proj{\ps}$) is however not even a valid event. In the classical
  example, we argued that each probability measure is uniquely
  determined by two actual coins. A similar (but much more subtle)
  argument is valid also in the quantum case. By postulates of quantum
  mechanics and Gleason's theorem, it turns out that for large enough
  quantum systems, each probability measure is uniquely determined by
  an actual quantum state as discussed next.
\end{example}

%%%%%
\subsection{Quantum Probability Measures}

Given our setup, the definition of a quantum probability measure is a
small variation on the classical definition.

\begin{definition}[Quantum Probability Measure~\cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010}]\label{def:QuantumProbabilitySpace}
Given a Hilbert space $\Hilb$ with its set of events~$\events$,
a \emph{quantum probability measure} is a function~$\mu:\events\rightarrow[0,1]$
such that: 
\begin{itemize}
\item $\mu(\mathbb{0})=0$. 
\item $\mu(\mathbb{1})=1$. 
\item For any projection $P$, $\mu\left(\mathbb{1}-P\right)=1-\mu\left(P\right)$.
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have $\mu\left(\sum_{i=1}^{N}P_{i}\right)=\sum_{i=1}^{N}\mu\left(P_{i}\right)$.
\end{itemize}
\end{definition}

\noindent A quantum probability measure can be easily constructed if
one knows the current state of the quantum system by using the Born
rule.  Specifically, for each
pure normalized quantum state $\ket{\phi}$, the Born rule induces a
probability measure $\mu_{\phi}^{\mathrm{B}}$ defined as
$\mu_{\phi}^{\mathrm{B}}(P)=\ip{\phi}{P\phi}$. The situation generalizes to
mixed states $\rho = \sum_{j=1}^{N}q_{j}\proj{\phi_{j}}$, where
$\sum_{j=1}^{N}q_{j}=1$ in which case the generalized Born rule
induces a probability measure $\mu_{\rho}^{\mathrm{B}}$ defined
as
$\mu_{\rho}^{\mathrm{B}}\left(P\right) = \Tr\left(\rho P\right) =
\sum_{j=1}^{N}
q_{j}\mu_{\phi_{j}}^{\mathrm{B}}\left(P\right)$~\cite{peres1995quantum,544199,Jaeger2007}.
Conversely every probability measure must be of this form.

\begin{thm}[Gleason's
  theorem~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}]\label{cor:Gleason's}In
a Hilbert space $\Hilb$ of dimension $d\geq3$, given a quantum probability
measure~$\mu:\events\rightarrow[0,1]$, there exists a unique mixed
state~$\rho$ such that $\mu=\mu_{\rho}^{\mathrm{B}}$.
\end{thm}

%%%%%
\subsection{Measuring Quantum Probabilities}

Similarly to the classical case, it is possible to estimate quantum
probabilities by utilizing the frequentist approach of the previous
section, assuming identical measurements conditions in each repeated
experiment~\cite{peres1995quantum}. For instance, if one wants to
determine the probability that the spin of a given silver atom is
$+\hbar/2$, a Stern-Gerlach apparatus is built where ideally an inhomogeneous
magnetic field is generated along, let's say, the quantization axis
$z$. One then produces a collimated beam of identically prepared
(neutral) silver atoms that is directed between the poles of the magnet
where a predetermined field-gradient along the $z$ direction has
been established. Under appropriate experimental conditions we will
observe that the beam, after traversing the magnetic-field region,
will be deflected towards two regions identified by distinguished
spots on a detector situated behind the apparatus~\cite{Stern1988,peres1995quantum,544199,Griffiths2003}.
Each of the two discrete values is associated to either $+\hbar/2$
or $-\hbar/2$, commonly called ``spin up'' and ``spin down'',
respectively. By ``counting'' the number of atoms that are deflected
in the ``spin up'' region one can, in principle, estimate the probability
that the prepared state of the silver atom state has spin $+\hbar/2$.
Notice that a real experiment does not necessarily represent an ``ideal
measurement''. For example, not all silver atoms will be identically
prepared, or the field-gradient could not be large enough to distinguish
between the spin up and down situations simply producing a large single
blot. In other words, the closer we get to an ideal measurement the
better we determine those probabilities at the cost of significantly
increasing the number of resources. It is not very well appreciated
in the literature that Bohr attempted to argue against the measurability
of the spin of a free electron~\cite{Bohr1985,MartensDeMuynck1994,McEvoy2001}.
Essentially, Bohr argued (and Mott later on justified his assertion
by an elegant use of uncertainty relations~\cite{10.2307/j.ctt7ztxn5.15})
that a Stern-Gerlach experiment could not succeed in establishing
the spin of an unbound electron because the Lorentz force would blur
the detected pattern. This example illustrates the case of a fundamental
physical limitation that not even infinite resources could mitigate.

\section{Cryptodeterminism}

>>>>

\begin{document}
\letter{Contextuality and the Born Rule in Fuzzy Quantum Theories}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The
Kochen-Specker theorem asserts that there is no function~$v$ assigning
a definitely value to all observables such that
\begin{equation}\eqalign{ 
v\left(O_{1}\right)+v\left(O_{2}\right)=v\left(O_{1}+O_{2}\right)\textrm{ and}\\
v\left(O_{1}\right)v\left(O_{2}\right)=v\left(O_{1}O_{2}\right)
}\label{eq:Kochen-Specker-Rules}\end{equation}
if $O_{1}$ and $O_{2}$ are commuting observables. Suppose that there
is a function~$v$ satisfying these two properties, it must send
an observable to its eigenvalues so that the following observables~\cite{Mermin1990Simple,peres1995quantum}:
\begin{equation}
\begin{array}{ccc}
\mathbb{1}\otimes\sigma_{z} & \sigma_{z}\otimes\mathbb{1} & \sigma_{z}\otimes\sigma_{z}\\
\sigma_{x}\otimes\mathbb{1} & \mathbb{1}\otimes\sigma_{x} & \sigma_{x}\otimes\sigma_{x}\\
\sigma_{x}\otimes\sigma_{z} & \sigma_{z}\otimes\sigma_{x} & \sigma_{y}\otimes\sigma_{y}
\end{array}\label{eq:MerminSquare}
\end{equation}
must be sent to $1$ or $-1$. In each row and column, operators commutes,
and each operator is the product of the two others, except in the
third column $\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clear impossible to find a function~$v$ such that
\begin{equation}
\begin{array}{ccc}
v\left(\mathbb{1}\otimes\sigma_{z}\right) & v\left(\sigma_{z}\otimes\mathbb{1}\right) & v\left(\sigma_{z}\otimes\sigma_{z}\right)\\
v\left(\sigma_{x}\otimes\mathbb{1}\right) & v\left(\mathbb{1}\otimes\sigma_{x}\right) & v\left(\sigma_{x}\otimes\sigma_{x}\right)\\
v\left(\sigma_{x}\otimes\sigma_{z}\right) & v\left(\sigma_{z}\otimes\sigma_{x}\right) & v\left(\sigma_{y}\otimes\sigma_{y}\right)
\end{array}\label{eq:MerminSquare-values}
\end{equation}
satisfying equation~(\ref{eq:Kochen-Specker-Rules}).
